---
title: "The cognitive basis of cooperative decision-making Experiments on uncertainty, frames, and time constraints"
author: "Nuno & Daniela"
date: '2022-08-31'
output: github_document
---
### Import Libraries
```{r, echo=FALSE,warning=FALSE, message=FALSE, results=FALSE}
packages <- c("astsa","fpp2","forecast","xts","TSstudio","ggplot2", "tseries", "MTS","dplyr","cowplot","stats","tidyr","egg","tidyverse","rlang","sjPlot","sjmisc","sjlabelled","readr","readxl","data.table","Matrix","lme4","emmeans")

installed_packages <- packages %in% row.names(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed.packages])
}

lapply(packages, library, character.only = TRUE)
```

# One-shot - First Experiment
### read excel
```{r}
data= read_excel("Firstexplong.xlsx")
```

## Data preprocessing
### In the BART TASK - 40551
```{r}

data$Investment <- factor(data$Investment, levels = c(0,1), labels = c("Not Invest", "Invest"))
data$Game<- factor(data$Game,levels = c(0,1), labels = c("SPD", "DPD"))
data$Pressure<- factor(data$Pressure,levels = c(0,1), labels = c("Time Delay", "Time Pressure"))   
data$Frame<- factor(data$Frame,levels = c(0,1), labels = c("Loss frame", "Gain Frame"))

data$Sex<- factor(data$Sex,levels = c(1,2), labels = c("Female", "Male")) 

data$svo<- factor(data$svo,levels = c(1,2), labels = c("Prosocial", "Proself"))

data$SES<- as.factor(data$SES)

```

## Exploratory data analysis
```{r}
#Table and proportions
table_1 <- table(data$Investment,data$Game)
table_2<-table(data$Investment,data$Pressure)
table_3<-table(data$Investment,data$Frame)
prop.table(table_1)
round(prop.table(table_1),2)

#row proportions, reference to row
round(prop.table(table_1,1),2)
100*round(prop.table(table_1,1),2)

#colum proportions
round(prop.table(table_1,2),2)

#table functions three-way table
table5 <-xtabs(~ Investment + Game + Pressure, data)
ftable(table5)
prop.table(table5)

DT <- data.table(data)
DT [, .("Cell N" = .N, 
       "Freq. of Cooperation" = sum(Investment == "Invest"), 
       "% Cooperating" = round(100 * sum(Investment == "Invest") / .N)),
   .(Frame, Game, Pressure)
   ]  %>% arrange(Frame, Game, Pressure)

as.data.table(data)
```

### Positive vs. Negative Frame
```{r}
DT[, .("Cell N" = .N, 
       "Freq. of Cooperation" = sum(Investment == "Invest"), 
       "% Cooperating" = round(100 * sum(Investment == "Invest") / .N)),
   .(Frame)
   ]  
```

### Type of game: DPD vs. SPD
```{r}
DT[, .("Cell N" = .N, 
       "Freq. of Cooperation" = sum(Investment == "Invest"), 
       "% Cooperating" = round(100 * sum(Investment == "Invest") / .N)),
   .(Game)
   ]  
```

### Time Pressure vs. Time Delay
```{r}
DT[, .("Cell N" = .N, 
       "Freq. of Cooperation" = sum(Investment == "Invest"), 
       "% Cooperating" = round(100 * sum(Investment == "Invest") / .N)),
   .(Pressure, Game)
   ]  
```


# Inferential statistics

##GLMM

## Full model - all variables + control
```{r}

model1=glm(Investment ~ Game*Frame*Pressure +svo*Pressure 
           +AdjustedValue_Bart*Game +AdjustedValue_Bart*Pressure +AdjustedValue_Bart*Frame
           +PerspectiveTaking_IRI +EmpathicConcern_IRI + PersonalDiscomfort_IRI 
           +Fantasy_IRI 
           +PANAS_Positive_Affect +	PANAS_Negative_Affect +Sex +SES +difficultyfelt 	
           +pleasant +proximity + cooperateonadailybasis
           ,family=binomial(link="logit"),data = data)

#summary(model1)
tab_model(model1)
logLik(model1)

```



## Model 2 - Manipulated variables + SES
```{r}
model2=glm(Investment ~ Game*Frame*Pressure
      +SES,
           family =binomial(link="logit"),data = data)

#summary(model2)
tab_model(model2)
logLik(model2)

```


### comparison across models 
```{r}
anova(model1,model2, test="LRT")
summary(anova(model1,model2))
        
```

## Time pressure manipulation check
```{r}

#sum(data$log10_rt >= 1 & data$Pressure==1, na.rm=TRUE)

#sum(data$log10_rt & data$Pressure ==1, na.rm=TRUE)
  

data_s <-  data %>% filter(log10_rt <= 1 & Pressure=="Time Pressure"|Pressure=="Time Delay")


m1=glm(Investment ~ Game*Frame*Pressure + svo*Pressure
           +AdjustedValue_Bart*Pressure+ AdjustedValue_Bart*Game+ AdjustedValue_Bart*Frame
          +PerspectiveTaking_IRI +EmpathicConcern_IRI + PersonalDiscomfort_IRI + Fantasy_IRI +PANAS_Positive_Affect +	PANAS_Negative_Affect
           + Sex + SES + difficultyfelt 	+pleasant
           +proximity + cooperateonadailybasis ,family = binomial(link="logit"),data = data_s)

#summary(m1)

tab_model(m1)
```

```{r}
m2=glm(Investment ~ Game*Frame*Pressure +
           +SES,family = binomial(link="logit"),data =data_s)

#summary(m2)

tab_model(m2)
```


# Experiment 2 - Repeated Games

### Read csv
```{r, echo=FALSE}
df = read.csv("dataSecondExp.csv")

names(df)[4] = "Game"
names(df)[1] = "Code"
names(df)[3] = "Frame"
names(df)[5] = "Pressure"
df$Code = factor(df$Code)
```

### Preprocess
```{r}

df$Investment <- factor(df$Investment, levels = c(0,1), labels = c("Not Invest", "Invest"))

df$Game<- factor(df$Game,levels = c(0,1), labels = c("SPD", "DPD"))
df$Pressure<- factor(df$Pressure,levels = c(0,1), labels = c("Time Delay", "Time Pressure"))   

df$Frame<- factor(df$Frame,levels = c(0,1), labels = c("Loss frame", "Gain Frame"))

df$Gender<- factor(df$Gender,levels = c(1,2), labels = c("Female", "Male")) 

df$svo_type[df$svo_type==4]<- 3

```

### remove subjects
```{r}
df_mod = subset(df, !(Code %in% c(43477,43509,
43725,43607)))

df_mod$Code = factor(df_mod$Code)
  

df_mod$svo_type[df_mod$svo_type==4]<- 3

```

### GLMM
```{r}


m1=glmer(Investment ~ Game*Frame*Pressure + (1|Supergame/Round/Code) +
             Gender + factor(SES)+ difficultyfelt+pleasant
             +proximity +cooperateonadailybasis 
             + factor(svo_type)*Pressure 
             + AdjustedValues_Bart*Pressure + AdjustedValues_Bart*Frame
             +AdjustedValues_Bart*Game
             +Perspective_Taking_IRI +Empathic_Concern_IRI
         +Personal_Discomfort_IRI
             +Fantasy_IRI
             +PANAS_Positive_Affect +PANAS_Negative_Affect 
             + Round + Supergame
             ,family = binomial,data = df_mod,
             control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
#summary(m1)

tab_model(m1)

```

### Marginal Means
```{r}
#warp = pairs(emmeans(m1 ~ Game*Frame*Pressure))
#warp

emm = emmeans(m1, ~Game*Frame*Pressure)
grid = ref_grid(m1)

emmip(grid,Pressure~Game|Frame) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"))

```

```{r}
noise <- emmeans(m1, ~Pressure*Game*Frame)


contrast(noise, "consec", simple = "each", combine = TRUE, adjust = "mvt")
```

### Other's previous decision
### preprocess
```{r}

Otherpastdecision= c()
contador = 10
counter = 1
for(i in 1:length(df_mod$Other.decision)){
     
     if (contador == 10){
         Otherpastdecision = append(Otherpastdecision, NA)}
     
     else {
       Otherpastdecision = append(Otherpastdecision,df_mod$Other.decision[i-1])}
     
     if (contador == 19){
         contador = 10}
     
     else{
         contador = contador+1}
  
  counter = counter+1
}


df_mod$Otherpastdecision = Otherpastdecision
#print(Otherpastdecision)

#sum(is.na(df$Otherpastdecision))

```

### Glmm with other's past decision
```{r}
m2=glmer(Investment ~ Game*Frame*Pressure + 
             Gender + factor(SES) +difficultyfelt
            +pleasant +proximity +cooperateonadailybasis 
             +Otherpastdecision + factor(svo_type)*Pressure 
             + AdjustedValues_Bart*Pressure + AdjustedValues_Bart*Frame +AdjustedValues_Bart*Game
             +Perspective_Taking_IRI +Empathic_Concern_IRI
             +Personal_Discomfort_IRI +Fantasy_IRI
             +PANAS_Positive_Affect +PANAS_Negative_Affect
             + Round + Supergame + (1|Supergame/Round/Code) 
             ,family=binomial,data=df_mod,control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

tab_model(m2)
#summary(m2)

```

### Glmm 
```{r}

m3=glmer(Investment ~ Game*Frame*Pressure + 
             Gender + factor(SES)+ difficultyfelt+pleasant
             +proximity +cooperateonadailybasis 
             + factor(svo_type)*Pressure 
             + AdjustedValues_Bart*Pressure + AdjustedValues_Bart*Frame +AdjustedValues_Bart*Game
             +Perspective_Taking_IRI +Empathic_Concern_IRI
         +Personal_Discomfort_IRI
             +Fantasy_IRI
             +PANAS_Positive_Affect +PANAS_Negative_Affect 
             + factor (Round) + factor (Supergame)  + (1|Supergame/Round/Code) 
             ,family = binomial,data = df,
             control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
#summary(m3)

tab_model(m3)

```

# Time-series Analysis

### Read csv
```{r, echo=FALSE}
df = read.csv("dataSecondExp.csv")
```

### Import Libraries
```{r, echo=FALSE,warning=FALSE, message=FALSE, results=FALSE}
packages <- c("astsa","fpp2","forecast","xts","TSstudio","ggplot2", "tseries", "MTS","dplyr","cowplot","stats","tidyr","egg")

installed_packages <- packages %in% row.names(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed.packages])
}

lapply(packages, library, character.only = TRUE)
```

### Filter for cooperative decisions
```{r}
df %>% filter(Investment == 1) %>% group_by(Supergame,Round) %>% summarize(mean = n()/192*100)->df2

df2$trial<-1:nrow(df2)

```
## Overall Results

### Decomposition of additive time series - trend, sazonality & noise

```{r}
df1 %>% decompose() %>% autoplot(range.bars = FALSE) + theme_bw() + xlab("Supergame") + scale_x_continuous(breaks=seq(1,10,1))+    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) +   theme(plot.title = element_text(hjust = 0.5))
```
### ACF & PACF plots
```{r}

acf(diff(ts), lag = 365)

pacf(diff(ts), lag = 365)
```

### Auto-Arima to the overall results
```{r}
arima = auto.arima(ts, trace=TRUE) 
```

### Ljung-Box test
```{r}
checkresiduals(arima)
```
### Arima summary
```{r}
summary(arima)
```

### Observations vs Sarima predictions - Plot
```{r}
df_ts = as.data.frame(ts)
df_arima = as.data.frame(fitted(arima))

df_bind = cbind(df_ts,df_arima)
names(df_bind)[1] = "Observations"
names(df_bind)[2] = "ARIMA"

df_bind %>% pivot_longer(cols=c('Observations', 'ARIMA'),
                    names_to='model',
                    values_to='mean') ->long

long$Round =  rep((10:109)/10, each=2)


plot_sarima = ggplot(long, aes(x=Round, y=mean, colour = factor(model), linetype = factor(model))) +
  geom_line(size = 0.7) + scale_linetype_manual(name = "Model", values = c(2,1),labels = c("SARIMA", "Observations")) +
  
  scale_color_manual(name = "Model", values = c("#F8766D","black"), labels = c("SARIMA", "Observations"))+
  
  scale_x_continuous(breaks=seq(1,10,1)) + scale_y_continuous(breaks=seq(0,100,10)) + 
  
  xlab("Supergame") + ylab("Mean Cooperation (%)") + theme_bw() +theme(plot.title = element_text(hjust = 0.5)) + theme(legend.title.align=0.5)+

    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + ggtitle("Observations vs Fitted Model")

plot_sarima

```

## Time Manipulation

## Pressure==1 (Time Pressure) 
```{r}
df %>% subset(Pressure == 1) -> df2
df2 %>% filter(Investment == 1) %>% group_by(Supergame,Round) %>% summarize(mean = n()/(nrow(df2)/100))->df2_pressure1

df2_pressure1$trial<-1:nrow(df2_pressure1)

```
### Auto-Arima
```{r}
ts <- ts(df2_pressure1$mean, start=1, frequency=10)
arima = auto.arima(ts, trace=TRUE) 
```

### Ljung-Box test
```{r}
checkresiduals(arima)
```
### Arima Summary
```{r}
summary(arima)
```

## Pressure==0 (Time-Delay)
```{r}
df %>% subset(Pressure == 0) -> df2
df2 %>% filter(Investment == 1) %>% group_by(Supergame,Round) %>% summarize(mean = n()/(nrow(df2)/100))->df2

df2$trial<-1:nrow(df2)

```

### Auto-Arima 
```{r}
ts <- ts(df2$mean, start=1, frequency=10)
arima = auto.arima(ts, trace=TRUE) 
```

### Arima Summary
```{r}
summary(arima)
```

### Ljung-Box
```{r}
checkresiduals(arima)
```

## Frame Manipulation

## Frame==1 (Gains)
```{r}
df %>% subset(Frame == 1) -> df2
df2 %>% filter(Investment == 1) %>% group_by(Supergame,Round) %>% summarize(mean = n()/(nrow(df2)/100))->df2

df2$trial<-1:nrow(df2)

```

### Auto-Arima
```{r}
ts <- ts(df2$mean, start=1, frequency=10)
arima = auto.arima(ts, trace=TRUE) 
```

### Arima Summary
```{r}
summary(arima)
```

### Ljung-Box
```{r}
checkresiduals(arima)
```

## Frame==0 (Losses)
```{r}
df %>% subset(Frame == 0) -> df2
df2 %>% filter(Investment == 1) %>% group_by(Supergame,Round) %>% summarize(mean = n()/(nrow(df2)/100))->df2

df2$trial<-1:nrow(df2)

```

### Auto-arima
```{r}
ts <- ts(df2$mean, start=1, frequency=10)
arima = auto.arima(ts, trace=TRUE) 
```

### Arima Summary
```{r}
summary(arima)
```

### Ljung-Box
```{r}
checkresiduals(arima)
```

## Game Manipulation

## Game==1 (Deterministic)
```{r}
df %>% subset(Game == 1) -> df2
df2 %>% filter(Investment == 1) %>% group_by(Supergame,Round) %>% summarize(mean = n()/(nrow(df2)/100))->df2

df2$trial<-1:nrow(df2)

```

### Auto-Arima
```{r}
ts <- ts(df2$mean, start=1, frequency=10)
arima = auto.arima(ts, trace=TRUE) 
```

### Arima Summary
```{r}
summary(arima)
```

### Ljung-Box
```{r}
checkresiduals(arima)
```

## Game==0 (Stochastic)
```{r}
df %>% subset(Game == 0) -> df2
df2 %>% filter(Investment == 1) %>% group_by(Supergame,Round) %>% summarize(mean = n()/(nrow(df2)/100))->df2

df2$trial<-1:nrow(df2)

```

### Auto-Arima
```{r}
ts <- ts(df2$mean, start=1, frequency=10)
arima = auto.arima(ts, trace=TRUE) 
```

### Arima Summary
```{r}
summary(arima)
```

### Ljung-Box
```{r}
checkresiduals(arima)
```

### Preprocess data for plotting

### Group-by Time condition
```{r}
df %>% filter(Investment == 1) %>% group_by(Supergame,Round,Pressure) %>% summarize(mean = n()/(nrow(df)/2/10000))->df2
df2$trial<-rep((10:109)/10, each=2)
```

```{r}
plot_time_pressure = ggplot(df2, aes(x=trial, y=mean, colour = factor(Pressure), linetype = factor(Pressure))) +
  geom_line(size = 0.7) + scale_linetype_manual(name = "Time", values = c(1, 6),labels = c("Time Delay", "Time Pressure")) +
  
  scale_color_manual(name = "Time", values = c("#00BFC4","#F8766D"), labels = c("Time Delay", "Time Pressure"))+
  
  scale_x_continuous(breaks=seq(1,10,1)) + scale_y_continuous(breaks=seq(0,100,10)) + 
  
  xlab("") + ylab("") + theme_bw() +theme(plot.title = element_text(hjust = 0.5)) + theme(legend.title.align=0.5)+

    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"))

#plot_time_pressure
```

### Group-by Frame condition
```{r}
df %>% filter(Investment == 1) %>% group_by(Supergame,Round,Frame.x) %>% summarize(mean = n()/(nrow(df)/2/10000))->df2
df2$trial<-rep((10:109)/10, each=2)
#head(df2,50)
```


```{r}
plot_frame = ggplot(df2, aes(x=trial, y=mean, color = factor(Frame.x), linetype = factor(Frame.x))) +
  geom_line(size = 0.7) + scale_linetype_manual(name = "Frame", values = c(1, 6),labels = c("Losses", "Gains")) +
  
  scale_color_manual(name = "Frame", values = c("#BB9D00","#AC88FF"), labels = c("Losses", "Gains"))+
  
  scale_x_continuous(breaks=seq(1,10,1)) + scale_y_continuous(breaks=seq(0,100,10)) + 
  
  xlab("") + ylab("Mean Cooperation (%)") + theme_bw() + theme(legend.title.align=0.5) +

    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"))

#plot_frame
```

### Group-by Game condition
```{r}
df %>% filter(Investment == 1) %>% group_by(Supergame,Round,Game) %>% summarize(mean = n()/(nrow(df)/2/10000))->df2
df2$trial<-rep((10:109)/10, each=2)
#head(df2,50)
```

```{r}
plot_game = ggplot(df2, aes(x=trial, y=mean, color = factor(Game), linetype = factor(Game))) +
  geom_line(size = 0.7) + scale_linetype_manual(name = "Game", values = c(1, 6),labels = c("Stochastic", "Deterministic")) +
  
  scale_color_manual(name = "Game", values = c("#00BC59", "#F066EA"), labels = c("Stochastic", "Deterministic"))+
  
  scale_x_continuous(breaks=seq(1,10,1)) + scale_y_continuous(breaks=seq(0,100,10)) + 
  
  xlab("Supergame") + ylab("") + theme_bw() + theme(legend.title.align=0.5) +

    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"))

#plot_game
```


# Plot Cooperation by Condition
```{r}
library(egg)


#same.size.ggplot(c("plot_time_pressure", "plot_frame", "plot_game"), "plot_frame") 
plot_row = plot_grid(plot_time_pressure, plot_frame, plot_game, nrow = 3, align = "v")

title <- ggdraw() + 
  draw_label(
    "Cooperation by Condition",
    fontface = 'bold',
    x = 0.25,
    hjust = 0
  ) +
  theme(
    # add margin on the left of the drawing canvas,
    # so title is aligned with left edge of first plot
    plot.margin = margin(0, 0, 0, 7)
  )
plot_grid(
  title, plot_row,
  ncol = 1,
  # rel_heights values control vertical title margins
  rel_heights = c(0.1, 1)
)
```

### Interaction Plot - Time*Game

### search for play with 0 cooperation
```{r}
counter = 1
contador=1

lista = c(0,1,10,11)

for (i in df2$new){
  if (i!=lista[counter]){
    print(contador)
    break
  }
  if (counter == 4){
    counter = 1
  }
  else {
    counter = counter+1
  }
  
  contador = contador+1
} 
```

```{r}

new_row = c(3,10,10,0)
df2 = as.data.frame(df2)
```

```{r}
df3 = rbind(df2[1:118,],new_row)
df4 = rbind(df3, df2[119:nrow(df2),])
```

```{r}
df4$trial<-rep((10:109)/10, each=4)
```

### recode
```{r}
df %>% 
       mutate(new = case_when(Game == 1 & Pressure == 1 ~ 11, 
                             Game == 1 & Pressure == 0 ~ 10,
                             Game == 0 & Pressure == 0 ~ 00,
                             Game == 0 & Pressure == 1 ~ 01)) -> df_new

```

### summarize
```{r}
df_new %>% filter(Investment == 1) %>% group_by(Supergame,Round,new) %>% summarize(mean = n()/(nrow(df_new)/2/10000))->df2
df2$trial<-rep((10:109)/10, each=4)
#head(df2,50)
```
#facet_wrap
```{r}


df2 %>% mutate(Game = case_when(new == 0 ~  0,
                                new == 1 ~  0,
                                new == 11 ~ 1,
                                new == 10 ~ 1),
               
               Pressure = case_when(new == 0 ~ 0,
                                 new == 1 ~ 1,
                                 new == 11 ~ 1,
                                 new == 10 ~0)) ->df5
```

### plot
```{r}
df5 %>%  mutate(Game = recode(Game, "0" = "Stochastic", "1" = "Deterministic")) %>% ggplot(aes(x=trial, y=mean, colour = factor(Pressure), linetype = factor(Pressure))) +
  geom_line(size = 0.7) + scale_linetype_manual(name = "Pressure", values = c(1, 6),labels = c("Time Delay", "Time Pressure")) +
  
  scale_color_manual(name = "Pressure", values = c("#00BFC4","#F8766D"), labels = c("Time Delay", "Time Pressure"))+
  
  scale_x_continuous(breaks=seq(1,10,1)) + scale_y_continuous(breaks=seq(0,100,10)) + 
  
  xlab("") + ylab("") + theme_bw() +theme(plot.title = element_text(hjust = 0.5)) + theme(legend.title.align=0.5)+

    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + facet_wrap(~Game)

```

